---
title: "p8105_hw3_yl5227"
author: "Jennifer Li"
date: "2023-10-12"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

Load `Instacart` data

```{r}
library(p8105.datasets)
data("instacart")

instacart = as_tibble(instacart)
```

*Short description of the dataset*

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. It encompasses a range of crucial variables, including customer identifiers, product information, such as _order id_, _user id_, _product name_. 

*How many aisles are there, and which aisles are the most items ordered from?*

```{r}
instacart |> 
  select(aisle) |> 
  distinct() |> 
  count()

instacart |> 
  count(aisle)  |> 
  arrange(desc(n)) |> 
  pull(aisle) |> 
  first() |> 
  print()
```

There are `r instacart |> select(aisle) |> distinct() |> count()` aisles, and the most items ordered from `r instacart |> count(aisle)  |> arrange(desc(n)) |> pull(aisle) |> first()` aisle.

**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The plot shows the number of items ordered in aisle with more than 10000 items ordered in an increasing order. The top two most popular aisle are fresh vegetables and fresh fruits.

**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  select(product_name, aisle) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(aisle, rank) |> 
  knitr::kable()
```

This table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits` from the most to the least popular items among the three most popular items.

**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)**

```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  select(order_hour_of_day, product_name, order_dow) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 2)
```

This table shows mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.


# Problem 2

Load `BRFSS` data.

```{r}
data("brfss_smart2010")

brfss = as_tibble(brfss_smart2010)
```


Clean this dataset

*   Use `janitor::clean_names()` to clean the variable names

*   Use `filter()` to subset the data of "Overall Health" topic

*   Order `response` variable from _Poor_ to _Excellent_

```{r}
brfss =
  brfss |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```

There are a total of `r length(brfss)` variables. There are `r nrow(brfss)` rows and `r ncol(brfss)` columns.

*In 2002, which states were observed at 7 or more locations? What about in 2010?*

```{r}
state_location_2002 =
  brfss |> 
  filter(year == 2002) |> 
  group_by(locationabbr) |> 
  distinct(locationdesc) |> 
  count(locationabbr) |> 
  filter(n >= 7) |> 
  print()

state_location_2010 =
  brfss |> 
  filter(year == 2010) |> 
  group_by(locationabbr) |> 
  distinct(locationdesc) |> 
  count(locationabbr) |> 
  filter(n >= 7) |> 
  print()
```

In 2002, CT, FL, MA, NC, NJ and PA were observed 7 or more locations. In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA were observed at 7 or more locations. 

*Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).*

The `brfss_excel` dataset is limited _Excellent_ responses, and contains year, state, and a variable that averages the data_value across locations within a state.

```{r}
brfss_excel =
  brfss |> 
  filter(response == "Excellent") |> 
  group_by(year, locationabbr) |> 
  summarize(avg_data_value = mean(data_value, na.rm = TRUE))

brfss_excel             
```

There are a total of `r length(brfss_excel)` variables. There are `r nrow(brfss_excel)` rows and `r ncol(brfss_excel)` columns.

Create the plot for `brfss_excel`.

```{r}
ggplot(brfss_excel, 
       aes(x = year, y = avg_data_value, 
           group = locationabbr, 
           color = locationabbr)) +
  geom_point() + 
  geom_line() +
  labs(
    title = "Average Data Value Over Time within a State",
    x = "Year",
    y = "Average Data Value",
    color = "State"
  ) +
  theme_minimal()
```

This plot shows the average data value over time within each state, limited to "Excellent" response. There are a total of `r length(unique(pull(brfss_excel, locationabbr)))` states observed. To make the plot more readable, the legends indicating `state` have been placed on the side of the plot, with the x-axis representing the `year`, and the y-axis representing the `average data value`. The `average data value` fluctuates by `year` for most states.

*Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.*

```{r}
ny_2006_2010 =
  brfss |>
  filter(locationabbr == "NY", 
         year %in% c(2006, 2010)) |>
  separate(locationdesc, into = c("state", "county"), "-") |>
  select(year, county, response, data_value) |>
  ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  geom_point(aes(color = county)) +
  facet_grid(~ year) +
  labs(title = "Data Value distribution in Responses among NY locations (in 2006 and 2010)",
       x = "Response",
       y = "Data Value",
       color = "County") +
  theme_minimal()

ny_2006_2010
```

`ny_2006_2010` is a plot that contains `brfss` that filters out only New York and years _2006_ or _2010_ data values. The above is a two panel plot showing, for the years 2006 and 2010, the distribution of the _data_value_ for responses (“Poor” to “Excellent”) among locations in New York State. The plot shows smililar distribution of _data_value_ for responses among locations in NY locations. In `2006`, `Good` and `Very good` responses are high in terms of mean `data value`. In `2010`, `Very good` response has the highest mean `data value`.

# Problem 3

Import dataset `nhanes_accel.csv` and `nhanes_covar.csv`, and clean variable names of these datasets.

```{r}
df_nhanes_accel =
  read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names()

df_nhanes_covar =
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names()
```

Data wrangling of `df_nhanes_covar` and `df_nhanes_accel`. 

*   Recode sex and education.

*   Change sex and education to factor variables. 

```{r}
df_nhanes_covar =
df_nhanes_covar |> 
  mutate(sex = case_match( sex,
        1 ~ "Male",
        2 ~ "Female"),
    education = case_match(education,
        1 ~ "less than high school",
        2 ~ "high school equivalent",
        3 ~ "more than high school"),
    education = as.factor(education),
    sex = as.factor(sex),)

df_nhanes_accel =
  df_nhanes_accel |> 
  pivot_longer(
    min1:min1440,
    names_prefix = "min",
    names_to = "time_minute",
    values_to = "mims_value")
```

Merge `df_nhanes_covar` and `df_nhanes_accel` and clean merged dataset. Filter out participants that are under 21 and exclude NAs.

```{r}
df_merged =
  merge(df_nhanes_covar, df_nhanes_accel, by = "seqn") |> 
  filter(age >= 21) |> 
  drop_na()
```

The `df_merged` merges `df_nhanes_covar` and `df_nhanes_accel` dataframes. It has `r nrow(df_merged)` rows and `r ncol(df_merged)` columns. 

*Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.*

```{r}
sex_ed =
  df_merged |> 
  select(sex, education, seqn) |>   
  group_by(sex, education) |> 
  distinct(seqn) |> 
  summarize(freq = n(), .groups = "drop") |> 
  pivot_wider(names_from = sex, values_from = freq) |> 
  knitr::kable()

print(sex_ed)
```

The above table `sex_ed` shows the number of men and women in each education category.The table shows that there's basically no gender difference within the same education level. Also, more people had the education with `more than high school` than in `high school equivalent` and `less than high school`.


Visualization of the age distributions for men and women in each education category

```{r}
age_distribution = 
  df_merged |> 
  ggplot(aes(x = age, fill = sex)) +
  geom_histogram(position = "identity", alpha = 0.7, bins = 5) +
  facet_grid(sex ~ education) +
  labs(
    title = "Age Distributions for men and women in each education category",
    x = "Age",
    y = "Frequency"
  ) +
  theme_minimal()

print(age_distribution)
```

The dataset highlights gender-specific patterns in education and age. Females slightly outnumber males with less than a high school education, while more males than females have a high school equivalent education. For those with education exceeding high school, there are slightly more females than males. Age-wise, men with less than high school education skew younger, high school equivalent education is bimodal, and education exceeding high school is uniformly distributed. Among women, less than high school education exhibits a uniform age distribution, high school equivalent education is left-skewed, and education exceeding high school is right-skewed, indicating varying age concentrations.


*Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.*

```{r}

```


*Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.*

```{r}

```

