---
title: "p8105_hw3_yl5227"
author: "Jennifer Li"
date: "2023-10-12"
output: github_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

Load `Instacart` data

```{r}
library(p8105.datasets)
data("instacart")

instacart = as_tibble(instacart)
```

## short description of the dataset

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. It encompasses a range of crucial variables, including customer identifiers, product information, such as _order id_, _user id_, _product name_. 

## How many aisles are there, and which aisles are the most items ordered from?

```{r}
instacart |> 
  select(aisle) |> 
  distinct() |> 
  count()

instacart |> 
  count(aisle)  |> 
  arrange(desc(n)) |> 
  pull(aisle) |> 
  first() |> 
  print()
```

There are `r instacart |> select(aisle) |> distinct() |> count()` aisles, and the most items ordered from `r instacart |> count(aisle)  |> arrange(desc(n)) |> pull(aisle) |> first()` aisle.

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

The plot shows the number of items ordered in aisle with more than 10000 items ordered in an increasing order. The top two most popular aisle are fresh vegetables and fresh fruits.

## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |> 
  select(product_name, aisle) |> 
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(aisle, rank) |> 
  knitr::kable()
```

This table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits` from the most to the least popular items among the three most popular items.

## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  select(order_hour_of_day, product_name, order_dow) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hour = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) |> 
  knitr::kable(digits = 2)
```

This table shows mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.


# Problem 2

Load `BRFSS` data.

```{r}
data("brfss_smart2010")

brfss = as_tibble(brfss_smart2010)
```


Clean this dataset

*   Use `janitor::clean_names()` to clean the variable names

*   Use `filter()` to subset the data of "Overall Health" topic

*   Order `response` variable from _Poor_ to _Excellent_

```{r}
brfss =
  brfss |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health") |> 
  mutate(response = fct_relevel(response, "Poor", "Fair", "Good", "Very good", "Excellent"))
```

